---
title: 'Peer Assesment2: U.S. Natural Dissasters 1950-2010'
author: "Sotero"
date: "Saturday, November 15, 2014"
output: html_document
---

#Sypnosys

Extreme weather events are a big problem in the U.S. In this article we will examine which event is the most deadly to Americans and the most costly. We have data from NOAA storm database. Our data consists of most natural disaster that have occurred in the U.S. from 1950-2010.    We first look at which natural disaster has caused the most fatalities and the most injuries. Second, we look at which natural disaster has been the most costly. 


# Data Processing


First we download the data.

```{r, cache=TRUE}

download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2","C:/Users/Sotero/Documents/AWC Laptop/RR/Week 3/PeerAssesment2/RRpeerassesment2/data.csv" )

data <-read.csv("data.csv", header= T )


head(data)

str(data)


```

In this part we find the total number of fatalities for each type of disaster.


```{r,cache=TRUE}



fatal  <- tapply(data$FATALITIES , data$EVTYPE , sum, na.rm= T)


fataldf <- data.frame(EVTYPE = names(fatal) , FATALITIES = fatal     )



```

No we find the disaster type with the most number of fatalities. 

```{r,cache = TRUE}

index <- which.max(fataldf$FATALITIES)

fataldf$EVTYPE[index]



```
Tornadoes killed a total of

```{r,cache=TRUE}

max(fataldf$FATALITIES)


```
people.

Now we will look for the disaster type that causes the most injuries

```{r,cache=TRUE}


injur <- tapply(data$INJURIES,data$EVTYPE,sum, na.rm = T  )

injurdf <- data.frame(EVTYPE = names(injur),INJURIES = injur   )

indexi <- which.max(injurdf$INJURIES)

injurdf$EVTYPE[indexi]



```


Tornadoes have caused

```{r,cache=TRUE}

max(injurdf$INJURIES)


```
injuries

Plots

```{r}





data$BGN_DATE <-strptime(data$BGN_DATE, format = "%m/%d/%Y %H:%M:%S" )
  
tordata <- subset(data,EVTYPE == "TORNADO"  )



# here we compute the total tornado fatalities per bgn_date. To do this we extrat the year from the #bgn_date

tordata$BGN_DATE <- format(tordata$BGN_DATE,"%Y")




toryear <- tapply(tordata$FATALITIES,tordata$BGN_DATE,sum)  

toryeardf <- data.frame(BGN_DATE = names(toryear),FATALITIES = toryear   )


toryeardf$BGN_DATE <- as.character(toryeardf$BGN_DATE )

toryeardf$BGN_DATE <-as.numeric(toryeardf$BGN_DATE )


plot(toryeardf$BGN_DATE , toryeardf$FATALITIES,xlab="Year",ylab="Fatalities",type="l",main=  "Tornados")

#dev.copy(png,file="tornadofatal.png")

#dev.off()

```


# Now examine any financial damage


```

In this part we find the total number of property damages for each type of disaster.


```{r,cache=TRUE}


dame  <- tapply(data$PROPDMG , data$EVTYPE , sum, na.rm= T)


damagesdf <- data.frame(EVTYPE = names(dame) , PROPDMG = dame     )



```

Now we find the disaster type with the most property damages

```{r,cache = TRUE}

indexd <- which.max(damagesdf$PROPDMG)

damagesdf$EVTYPE[indexd]



```
Tornados have damaged

```{r,cache=TRUE}

max(damagesdf$PROPDMG)


```
properties.


```{r}

# here we find the total number of properties damaged each year by tornados

dmy <- tapply(tordata$PROPDMG,tordata$BGN_DATE,sum)  

damprop <- data.frame(BGN_DATE = names(dmy),PROPDMG = dmy   )


damprop$BGN_DATE <- as.character(damprop$BGN_DATE )

damprop$BGN_DATE <-as.numeric(damprop$BGN_DATE )


plot(damprop$BGN_DATE,damprop$PROPDMG,type="l",xlab="Year",ylab = "Property Damage", main= "Tornado"  )



```


```{r,cache=TRUE}

data$PROPDMGEXP <- as.numeric(as.character(data$PROPDMGEXP))



# mydata$multiple <- 1
# mydata$multiple[mydata$PROPDMGEXP =="K"] <- 1000
# mydata$multiple[mydata$PROPDMGEXP =="M"] <- 1000000

data$PROPDMGEXP[data$PROPDMGEXP=="K"]  <- 1000
data$PROPDMGEXP[data$PROPDMGEXP=="M"] <- 1000000
data$PROPDMGEXP[data$PROPDMGEXP=="B"] <- 1000000000

data$damages <- data$PROPDMGEXP*data$PROPDMGEXP





```


# Results

Our data shows that tornados are the most deadly, harmful and costly natural disaster in the United States.  Tornados have caused a total of 5633 fatalities and 91346 injuries on the time period from 1950-2010.  In addition to the total damages by tornadoes is 3212258. No other natural disaster has caused as many fatalities, injuries, and damages. The graph shows that fatalities due to tornarnadoes have not increased with the exception of 2010. Howver, damages due to tornadoes have increased over the years. This is show on the second figure.



