---
title: 'Peer Assesment2: Natural Dissasters'
author: "Sotero"
date: "Saturday, November 15, 2014"
output: html_document
---

# Data Processing


First we download the data.

```{r, cache=TRUE}

download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2","C:/Users/Sotero/Documents/AWC Laptop/RR/Week 3/PeerAssesment2/RRpeerassesment2/data.csv" )

data <-read.csv("data.csv", header= T )




```

In this part we find the total number of fatalities for each type of disaster.


```{r,cache=TRUE}

str(data)

head(data)


fatal  <- tapply(data$FATALITIES , data$EVTYPE , sum, na.rm= T)


fataldf <- data.frame(EVTYPE = names(fatal) , FATALITIES = fatal     )



```

No we find the disaster type with the most number of fatalities. 

```{r,cache = TRUE}

index <- which.max(fataldf$FATALITIES)

fataldf$EVTYPE[index]



```
Tornadoes killed a total of

```{r,cache=TRUE}

max(fataldf$FATALITIES)


```
people.

Now we will look for the disaster type that causes the most injuries

```{r,cache=TRUE}


injur <- tapply(data$INJURIES,data$EVTYPE,sum, na.rm = T  )

injurdf <- data.frame(EVTYPE = names(injur),INJURIES = injur   )

indexi <- which.max(injurdf$INJURIES)

injurdf$EVTYPE[indexi]



```


Tornadoes have caused

```{r,cache=TRUE}

max(injurdf$INJURIES)


```
injuries

Plots

```{r,cache=TRUE}

library("ggplot2", lib.loc="~/R/win-library/3.1")





data$BGN_DATE <-strptime(data$BGN_DATE, format = "%m/%d/%Y %H:%M:%S" )
  
tordata <- subset(data,EVTYPE == "TORNADO"  )



# here we compute the total tornado fatalities per bgn_date. To do this we extrat the year from the #bgn_date

tordata$BGN_DATE <- format(tordata$BGN_DATE,"%Y")




toryear <- tapply(tordata$FATALITIES,tordata$BGN_DATE,sum)  

toryeardf <- data.frame(BGN_DATE = names(toryear),FATALITIES = toryear   )


toryeardf$BGN_DATE <- as.character(toryeardf$BGN_DATE )

toryeardf$BGN_DATE <-as.numeric(toryeardf$BGN_DATE )


#plot(toryeardf$BGN_DATE , toryeardf$FATALITIES,xlab="Year",ylab="Fatalities",type="l",main=  "Tornados")

#dev.copy(png,file="tornadofatal.png")

#dev.off()

```


# Now examine any financial damage


```

In this part we find the total number of property damages for each type of disaster.


```{r,cache=TRUE}


dame  <- tapply(data$PROPDMG , data$EVTYPE , sum, na.rm= T)


damagesdf <- data.frame(EVTYPE = names(dame) , PROPDMG = dame     )



```

Now we find the disaster type with the most property damages

```{r,cache = TRUE}

indexd <- which.max(damagesdf$PROPDMG)

damagesdf$EVTYPE[indexd]



```
Tornados have damaged

```{r,cache=TRUE}

max(damagesdf$PROPDMG)


```
properties.


```{r,cache=TRUE}

# here we find the total number of properties damaged each year by tornados

dmy <- tapply(tordata$PROPDMG,tordata$BGN_DATE,sum)  

damprop <- data.frame(BGN_DATE = names(dmy),PROPDMG = dmy   )


damprop$BGN_DATE <- as.character(damprop$BGN_DATE )

damprop$BGN_DATE <-as.numeric(damprop$BGN_DATE )


#plot(damprop$BGN_DATE,damprop$PROPDMG  )

#with(data=damprop,plot(BGN_DATE,PROPDMG ) )



```



